# Basics

In this chapter, basic elements needed for using R will be inspected.


## Vectors, matrices and lists
- **Vector**  
vector can contain only same class of objects
```{r}
d <- c(1:5)
d[d %in% 1:3] <- 0
d
```


- Sequence `seq`
```{r}
1:14 # this can be done also with the `seq` as below
seq(1, 14) # this gives all the number from 1 to 14
seq(1, 14, by = 2) # this gives odd numbers within the range of 1 to 14
seq(2, 14, by = 2) # this gives even numbers within the range of 2 to 14
k <- seq(from = 100, to = 1000, by = 20)
str(k)
```

- Replication `rep`
```{r}
x <- rep(c(1:2),5)
x

```



- **Vectorized operations**  
see how the data parallels
```{r}
e <- 1:4; f <-6:9
e + f # 1+6, 2+7...
e*f # 1*6, 2*7
```


- **Matrix**  
```{r}
x <- c(1:9)
matrix1 <- matrix(data = x, nrow = 3, ncol = 3)
matrixbyrow <- matrix(data = x, nrow = 3, ncol = 3, byrow = TRUE) # by row
matrix1
matrixbyrow
```


```{r}
int <- 1:16
matrix2 <- matrix(int, nrow = 4, ncol = 4)
matrix2[3,3] #the value in 3rd row and 3rd column of the matrix2
matrix2[3,] # all the values in 3rd row
matrix2[,3] # all the values in 3rd column
matrix2[1:3,1:3] # all the values first 3rd rows and 3rd columns
matrix2[-4, -4] # this gives the same outcome as the one above
matrix2[c(1,3),] # values in 1st and 3rd rows
matrix2[,c(1,3)] # values in 1st and 3rd columns
```


- **Matrices multiplication**  
`;` and `%*%`
```{r}
g <- matrix(1:4,2,2); h <- matrix(seq(10,40, by=10),2,2) # four 10s 2 by 2
g; h
g*h # 1*10, 2*20, 3*30, 4*40
g%*%h # (1*10)+(3*20), (2*10)+(4*20), (1*30)+(3*40), (2*30)+(4*40)
```

- **Lists**  
store different kinds of class of objects
```{r}
a <- c("A","B","C","D","E")
b <- seq(0, 1, 0.1)
c <- c(T,F,F,T,T,F,T,T,F)

list <- list(a, b, c)
list
list[[2]] # "b"
list[[2]][1:3] # 1:3 values in "b"
```


- **Names**  
`paste`
```{r}
z <- 1:3
paste(z, collapse = " ")
```

 `names`
```{r}
names(z) <- c("Y", "U", "I")
z
```

`dimnames`
```{r}
m <- matrix(1:6,3,2)
dimnames(m) <- list(c("a","b","c"), c("A","B"))
m
```


- **Levels**
```{r}
fac <- factor(c("1 star", "2 stars", "3 stars", "3 stars", "4 stars", "5 stars"))
levels(fac) # what values there are in the vector
nlevels(fac) # how many levels there are

fac[5] # 5th value in the vector (which is 4 stars in this case) 
fac[2:5] # 2nd to 5th values in the vector
fac[-1] # except for the 1st value
```



- **samples**  
```{r}
set.seed(1)
sample(1:10,4)
sample(letters,5)
breakfast <- c("coffee","tea","milk")
table(sample(breakfast, 50, replace = T, prob = c(0.4,0.4,0.2)))
```


