# Functions



## Lexical scoping
- **using operator `<<-`**  
which assign a value to an object in global environment
```{r}
makeVector <- function(x = numeric()) {
        m <- NULL # create a space which we can use
        set <- function(y) {
                x <<- y # put y into global environment
                m <<- NULL # makes it possible to clear previous values
        }
        get <- function() x
        setmean <- function(mean) m <<- mean
        getmean <- function() m
        list(set = set, get = get,
             setmean = setmean,
             getmean = getmean) # makes is possible to retrieve results using $
}
```

`cachemean` is needed to complete `makeVector` as follows.
Note that cachemean() is the only place where the mean() function is executed, which is why makeVector() is incomplete without cachemean().
```{r}
cachemean <- function(x, ...) {
        m <- x$getmean()
        if(!is.null(m)) {
                message("getting cached data")
                return(m) # checks if there is any cached data
        }
        data <- x$get() # if not cached, it gets the value
        m <- mean(data, ...) # and calculate the mean
        x$setmean(m) # set it
        m # and return it :)
}
```


When you run the second line using `cachemean` for the first time, there appears no message, but when you run it for the second time, it says "getting cached data", meaning there is a data already stored.
```{r}
myVector <- makeVector(x = 1:15)
cachemean(myVector) 
```

