--- 
title: "A Useful Code Book"
author: "Yui Naruse"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Introduction

This is a code book written in **Markdown** using `bookdown` package developed by Yihui Xie.
The guide book can be found [here](https://bookdown.org/yihui/bookdown/).

The motivation behind writing this book is to make it easier to review the contents. Therefore, it will mainly contain basics of the R with minimum explanations of the codes accompanied by some examples. 
Reference Chapter with \@ref(intro).

Here are some of the `packages` I will be using in this book.
```{r message=FALSE}
library(devtools)
library(dplyr)
library(ggplot2)
library(statsr)
library(shiny)
library(tidyverse)
library(vcd)
library(ggpubr)
library(knitr)
library(kableExtra)
library(GGally)
library(UsingR)
library(bookdown)
Sys.setenv(LANG = "en")
```


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```



