---
title: "Important"
author: "Yui Naruse"
date: "2021/6/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

### R Programming

#### Functions
**Anonymous functions**
Just as the case below, although the function `evaluate` did not have function `sd` defined, we can use it in the function as anonymous function.
```{r}
evaluate <- function(func, dat){
  func(sum(dat))
  func(median(dat))
  func(floor(dat))
}

evaluate(sd, c(4, 8, 0))
evaluate(function(x){x+1}, 6)
evaluate(function(x){x[1]}, c(8, 4, 0))

```

**paste**
```{r}
telegram <- function(...){
  paste("START", ..., "STOP")
}
```


**ellipsis, args, list**
```{r}
mad_libs <- function(...){
  args <- list(...)
  place <- args[["place"]]
  adjective <- args[["adjective"]]
  noun <- args[["noun"]]
  paste("News from", place, "today where", adjective, "students took to the streets in protest of the new", noun, "being installed on campus.")
}
mad_libs(place = "my house", adjective = "boring", noun = "cat")
```

**binary operators**
```{r}
"%p%" <- function(left, right){
  paste(left, right)
}
"I" %p% "Love" %p% "You!"
```

**Prime numbers**
```{r}
isItPrime <- function(n) {
  n.int <- as.integer(n)
  if(!(n.int == n)) stop("n is not an integer")
  if(n < 1) stop("n is not positive")
  if(n.int == 1) return(FALSE)
  if(n.int == 2) return(TRUE)
  lastq <- as.integer(sqrt(n)) + 1L
  for (q in 2:lastq) {
    if((n %% q) == 0) return(FALSE)
  }
  return(TRUE)
}
isItPrime(8)

```


**for loop**
for assignment week 2. 
This might also be helpful to see how others are subsetting the data
(https://stackoverflow.com/questions/24154527/warning-message-in-mean-default-argument-is-not-numeric-or-logical-returning-na)
```{r}
polmean2 <- function(directory, pollutant, id = 1:332) {
  files_list <- list.files(directory, full.names = TRUE)
  dat <- data.frame()
  for (i in id) {
    dat <- rbind(dat, read.csv(files_list[i]))
  }
  if(pollutant == "sulfate") {
    pollutant <- dat[,2]
    return(mean(pollutant, na.rm = TRUE))
  } else {
    pollutant <- dat[,3]
    return(mean(pollutant, na.rm = TRUE))
  }
  }

polmean2("specdata", "sulfate", id = 30:70)

## Note that this can be done using dplyr as bellow
specdata %>%
  filter(ID %in% 30:70) %>%
  summarise(mean = mean(sulfate, na.rm = T))

```

**for loop 2**
for assignment week 2. 

```{r}
complete4 <- function(directory, id = 1:332) {
  files_list <- list.files(directory, full.names = TRUE)
  id_len <- length(id)
  complete_data <- rep(0, id_len) # somehow had to make a vector of 0s with length of ids
  j <- 1
  for (i in id) {
    dat <- read.csv(files_list[i])
    complete_data[j] <- sum(complete.cases(dat))
    j <- j + 1 # when i do not include this line, it gives only one value of specific id
  }
  result <- data.frame(id, nods = complete_data)
  return(result)
}

complete4("specdata", id = c(5,60,20))

```
**correlation function**


```{r}
corr1 <- function(directory, threshold = 0) {
  if(grep("specdata", directory) == 1) {
    directory <- ("./specdata/")
  }
  complete_table <- complete4("specdata", 1:332)
  nods <- complete_table$nods
  ids <- complete_table$id[nods > threshold]
  id_len <- length(ids)
  corr_vector <- rep(0, id_len)
  all_files <- as.character(list.files(directory))
  file_paths <- paste(directory, all_files, sep = "")
  j <- 1
  for (i in ids) {
      current_file <- read.csv(file_paths[i], header = T, sep = ",")
      corr_vector[j] <- cor(current_file$sulfate, current_file$nitrate, use = "complete.obs")
      j <- j + 1
  }
  result <- corr_vector
  return(result)
}
```


**caching data**
```{r}
makeCacheMatrix1 <- function(x = matrix()) {
  m <- NULL
  set <- function(y) {
    x <<- y
    m <<- NULL
  }
  get <- function() x
  setinverse <- function(inverse) m <<- inverse
  getinverse <- function() m
  list(set = set, get = get,
       setinverse = setinverse,
       getinverse = getinverse)
}
```

**getting value from cached data**
```{r}
cacheSolve1 <- function(x, ...) {
  m <- x$getinverse()
  if(!is.null(m)) {
    message("getting cached data")
    return(m)
  }
  data <- x$get()
  m <- solve(data, ...)
  x$setinverse(m)
  m
}


mat <- makeCacheMatrix1(x = matrix(c(3,5,7,12,14,16,21,25,29,33,38,62,84,72,98,100),4,4))
cacheSolve1(mat)
```


